---
title: "casestudy1"
author: "Malcolm Carlson"
date: "10/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(magrittr)
library(tidyr)
library(plyr)
library(dplyr)
library(tidyverse)


beerdf=read.csv("C:\\Users\\malco\\Documents\\MDS\\MDS_6306_DDS\\CaseStudy1\\Beers.csv",header=TRUE)
breweriesdf=read.csv("C:\\Users\\malco\\Documents\\MDS\\MDS_6306_DDS\\CaseStudy1\\Breweries.csv",header=TRUE)
#IBU - International Bitterness Units
#ABV - Alcohol by Volume

head(breweriesdf) 
#Brew_ID,Name,City,State
breweriesdf <- breweriesdf %>% rename(
  Brewery_id = Brew_ID,
  Brewery_name = Name
)
head(breweriesdf)

head(beerdf)
beerdf <- beerdf %>% rename(
  Beer_name = Name
)
head(beerdf)
```

```{r}
combineddf<-merge(breweriesdf,beerdf, by="Brewery_id",all = TRUE)

head(combineddf)
```
#1. How many breweries are present in each state?
```{r}
combineddf%>%group_by(State)%>%summarise(count=n())
#need a bar chart in descending order
combineddf%>%ggplot(aes(x=State,fill=State))+geom_bar()
```
#3. Address the missing values in each column.
```{r}

sum(is.na(combineddf$Brewery_id))
sum(is.na(combineddf$Brewery_name))
sum(is.na(combineddf$City))
sum(is.na(combineddf$State))
sum(is.na(combineddf$Beer_name))
sum(is.na(combineddf$Beer_ID))
sum(is.na(combineddf$Style))
sum(is.na(combineddf$Ounces))
sum(is.na(combineddf$ABV))
sum(is.na(combineddf$IBU))

str(combineddf)
head(combineddf,1)


```
#4. Compute the median alcohol content and international #bitterness unit for each state. Plot a bar chart to compare. ABV
```{r}
medianABVIBU<-combineddf%>%group_by(State)%>%summarize(medianABV=median(ABV),medianIBU=median(IBU))
medianABVIBU


```
#5. Which state has the maximum alcoholic (ABV) beer? Which #state has the most bitter (IBU) beer?
```{r}
sortedbyIBU=medianABVIBU[order(-medianABVIBU$medianIBU),]
sortedbyABV=medianABVIBU[order(-medianABVIBU$medianABV),]
head(sortedbyABV)
```
#6. Comment on the summary statistics and distribution of the #ABV variable.
```{r}


```
#7. Is there an apparent relationship between the bitterness #of the beer and its alcoholic content? Draw a scatter plot.  #Make your best judgment of a relationship and EXPLAIN your #answer.
```{r}


```
#8.  Use KNN - For final presentation
```{r}

```
#9.  Look at, have a direction!
